---
title: 'Life cycle variation and resource dynamics ABM: Code'
author: "Pablo J. Varas Enríquez"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Outline

The following code is meant to develop an agent-based model to understand how the resource dynamics of female individuals influence the variability of life cycles at the population level. 

At first, there is a section called "Starting point", which are the first attempts and checks of the model. Different modules are used to build the initial states of the model, with less life cycle stages and resource dynamics. Once less modules work, more modules will be developed and included in the model. The starting point has two life cycle stages (i.e. juvenile, reproductive career), and two resource dynamics (i.e. production, consumption).

The following sections are the different scenarios where the model will be used. The scenarios are:

- A population where all individuals have medium probabilities of resource production and sharing.
- A population where all individuals have high probabilities for resource production and sharing.
- A population where most individuals (75% of the population) have a high probability of production and low probability of sharing. The rest (25%) have medium probabilities for production and sharing.
- A population where most individuals (75% of the population) have a low probability of production and high probability of sharing. The rest (25%) have medium probabilities for production and sharing.
- A population where all individuals have low probabilities for resource production and sharing.

### Starting point

The starting point section consists of the different attempts to start building the model. First, there will be one life cycle stage and the resource dynamics of production and consumption. Second, one with two life cycle stages, production, and consumption. Third, storage will be added, and finally, one with sharing dynamics.

#### First attempt: one life cycle, two resource dynamics.

At first, a population of 1000 individuals will be used for the model. The following variables are recorded for each individual:

- `ID`: identity of the individual.
- `age`: number of years the individuals lives. Each iteration that the individual survives makes her one year older.
- `age_repro`: age at which the individual has its first descendant.
- `n_repro`: number of descendants.
- `prod`: total amount of resources produced by the individual in her life cycle.
- `cons`: total amount of resources consumed by the individual in her life cycle.

```{r}
#make dataframe with life history traits and resource dynamics
pop <- data.frame(id=1:1000,age=rep(NA,length.out=1000),age_repro=rep(NA,length.out=1000),n_repro=rep(NA,length.out=1000),prod=rep(NA,length.out=1000),cons=rep(NA,length.out=1000))
#check data frame
head(pop)
```

You set the auxiliary variables, which are fixed values set at initialisation to define the life-history and resource dynamics that individuals can experience. Regarding life-history dynamics, you define the surviving probability (`surv_p`) and cost (`surv_c`), the reproductive probability (`repro_p`), cost (`repro_c`), and number of descendants per reproduction (`repro_n`). Resource-related auxiliary variables are the probability of production (`prod_p`) and consumption (`cons_p`), with the respective amount of resources associated (`prod_n` and `cons_n`, respectively). An individual can only produce one descendant in case of a reproductive event.

```{r}
####values are made up for now

#surival
#probability of surviving
surv_p <- 0.75
#cost of surviving
surv_c <- 0.25

#reproduction
#probability
repro_p <- 0.5
#cost of reproduction
repro_c <- 0.25
#number of descendants per reproduction
repro_n <- 1

#production
#probability
prod_p <- 0.5
#amount of resources produced
prod_n <- 1

#consumption
#probability
cons <- 0.5
#amount of resources consumed
cons_n <- 0.25
```

You set the functions for the state variables related to resource dynamics (i.e. production and consumption). The outcomes of the functions can be positive (1) or negative (0) from binomial distributions that are based on the values of the auxiliary variables.

```{r}
#production
produce <- function(ind){
  ind$prod <- sample(x=rbinom(1,1,prod_p),size=nrow(ind),prob=prod_p)
  return(ind)
}

#consumption
consume <- function(ind){
  ind$cons <- sample(x=rbinom(1,1,cons_p),size=nrow(ind),prob=cons_p)
  return(ind)
}
```

You set the functions to define the amount of resources the individuals has for herself. If the individual has a positive outcome (1) for a resource dynamic it adds the fixed amount of resource set up earlier into her current amount of resources. Otherwise (0), the amount of resources remains equal.

```{r}
#getting resources from production
prod_res <- function(ind){
  if(ind$prod==1)
  {
    ind$res_prod <- ind$res_prod+prod_n
  }
  else{
    ind$res_prod <- ind$res_prod
  }
  return(ind)
}

#losing resources from consumption
cons_res <- function(ind){
  if(ind$cons==1)
  {
    ind$res_cons <- ind$res_cons+cons_n
  }
  else{
    ind$res_cons <- ind$res_cons
  }
  return(id)
}
```

You set the functions for survival and reproduction (i.e. state variables). The probabilities of each function include the probability of survival/reproduction and the resources available from the different dynamics. A positive outcome is a 1 and a negative one is 0.

```{r}
#survival
survive <- function(ind){
  ind$surv <- sample(x=rbinom(1,1,surv),size=nrow(ind),prob=surv*ind$res_prod*ind$res_cons)
  return(ind)
}

#reproduction
reproduce <- function(ind){
  ind$repro <- sample(x=rbinom(1,1,repro),size=nrow(ind),prob=repro*ind$res_prod*ind$res_cons)
  return(ind)
}
```

You set the functions to include the resource costs related to survival and reproduction. If the individual has a positive outcome (1) for either life-history trait then the fixed amount of resources are substracted from the amount of resources available. If the outcome is negative (0) then either the individual dies, in the case of survival, or does not produce a descendant, in the case of reproduction.


```{r}
#survival-resource dynamics
surv_res <- function(ind){
  if(ind$surv==1)
  {
    ind$res_prod <- ind$res_prod-surv_c
  }
  else{
    ind$res_prod <- ind$res_prod
  }
  return(ind)
}

#reproduction-resource dynamics
repro_res <- function(ind){
  if(ind$repro==1)
  {
    ind$res_prod <- ind$res_prod-repro_c
    ind$repro_n <- repro_n+1
  }
  else{
    ind$res_prod <- ind$res_prod
    ind$repro_n <- repro_n
  }
  return(id)
}
```




